---
title: "Implementing the Basic SIR Model"
author: "Your Name"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Implementing the Basic SIR Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

This vignette provides a detailed guide on how to implement the basic SIR model using the provided R code in this repository. The SIR model is a simple mathematical model used to describe the spread of infectious diseases in a population. The model divides the population into three compartments: Susceptible (S), Infected (I), and Recovered (R).

# Prerequisites

Before you begin, make sure you have the following packages installed:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r prerequisites}
install.packages(c("deSolve", "magrittr", "tidyverse"))
```

# Loading the Required Libraries

First, load the required libraries:

```{r load-libraries}
library(deSolve)
library(magrittr)
library(tidyverse)
```

# Setting Up the Working Directory

Set up the working directory to the root of the repository:

```{r set-working-directory}
setwd("~/path/to/your/repository")
```

# Sourcing the Functions

Source the SIR simulation functions from the `R/Functions_SIR.R` file:

```{r source-functions}
source("R/Functions_SIR.R")
```

# Defining Model Parameters

Define the model parameters, including population sizes, time increments, and the basic reproduction number (R0):

```{r define-parameters}
popsize1 <- 10000 # population 1 size
popsize2 <- 5000 # population 2 size
popsize3 <- 9000 # population 3 size
dt <- 1  # time increment for simulation output (days)
time.vals <- seq(0,800,7) # time intervals

r0 <- 1.4 # basic reproduction number

# Set parameter vector
theta <- c(beta = NA,
          gamma = 1/10, # mean duration of infectiousness
          alpha1 = 0.0001, # relative contact rate between populations 1-2
          alpha2 = 0.0001, # relative contact rate between populations 2-3
          alpha3 = 0.000, # relative contact rate between populations 3-1
          npop1 = popsize1, # size of population 1
          npop2 = popsize2, # size of population 2
          npop3 = popsize3, # size of population 3
          rep = 0.25 # proportion of cases reported
          )

theta[["beta"]] <- r0*theta[["gamma"]] # define beta as function of R0 and gamma
```

# Setting Initial Conditions

Set the initial conditions for the simulation, assuming everyone is susceptible except for a few initial infections:

```{r set-initial-conditions}
initial_inf <- 1 # initially infectious in population 1
theta_init <- c(s_init=NA,i1_init=initial_inf,r_init=0,s2_init=NA,i2_init=0,r2_init=0,s3_init=NA,i3_init=0,r3_init=0)
theta_init[["s_init"]] <- theta[["npop1"]]-theta_init[["i1_init"]]
theta_init[["s2_init"]] <- theta[["npop2"]]-theta_init[["i2_init"]]
theta_init[["s3_init"]] <- theta[["npop3"]]-theta_init[["i3_init"]]
```

# Running the Simulation

Run the simulation using the defined parameters and initial conditions:

```{r run-simulation}
output_sim <- run_simulation(dt,theta,theta_init,time.vals)
```

# Plotting the Results

Plot the new reported cases over time for each population:

```{r plot-results}
par(mfrow=c(1,1),mar = c(3.2,3.5,1,1),mgp=c(2.2,0.6,0),las=1)

# Total cases
plot(time.vals/7,output_sim$C1_trace,type="l",col='white',xlab="weeks",ylab="total cases",ylim=c(0,500))
lines(time.vals/7, output_sim$C1_trace ,col='blue') # population 1
lines(time.vals/7, output_sim$C2_trace,col='red') # population 2
lines(time.vals/7, output_sim$C3_trace,col='dark green') # population 3
lines(time.vals/7, output_sim$C1_trace + output_sim$C2_trace + output_sim$C3_trace ,col='black',lwd=2)  # total cases

# Reported cases
plot(time.vals/7, output_sim$C1_trace,type="l",col='white',xlab="weeks",ylab="reported cases",ylim=c(0,120))
lines(time.vals/7, cases_reported(0.8*output_sim$C1_trace,theta) ,col='blue') # population 1
lines(time.vals/7, cases_reported(0.8*output_sim$C2_trace,theta),col='red') # population 2
lines(time.vals/7, cases_reported(0.8*output_sim$C3_trace,theta),col='dark green') # population 3
lines(time.vals/7, cases_reported(1.3*( output_sim$C1_trace + output_sim$C2_trace + output_sim$C3_trace),theta) ,col='black',lwd=2)  # total cases
```

# Saving the Plots

Save the plots as a PDF file:

```{r save-plots}
dev.copy(pdf,paste("plots/SIR_model.pdf",sep=""),width=8,height=4)
dev.off()
```

# Conclusion

This vignette provided a step-by-step guide on how to implement the basic SIR model using the provided R code in this repository. By following these steps, you can simulate the spread of infectious diseases in a population and visualize the results.
